Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-23 paper.qmd
Map the small cluster color to projections

scaled_mnist_data <- scaled_mnist |>
  filter(type == "data") |>
  select(-type)

@@

# Combine with the true model for visualization
df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
                       scaled_mnist_data |> mutate(type = "data"))

@@

@@

mnist_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.2,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b")

@@

mnist_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.2,
    alpha = 0.1) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c")

@@

mnist_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.2,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("d")

>>>>

scaled_mnist_data <- scaled_mnist |>
  filter(type == "data") |>
  # bind_cols(tsne_minst_scaled_temp) |>
  # select(-type, -tSNE1, -tSNE2, -ID) |>
  select(-type)

# scaled_mnist_data <- bind_cols(scaled_mnist_data,
#                                tsne_minst_scaled |> select(-ID))

@@

# model_2d <- df_bin_centroids_mnist |>
#   select(c_x, c_y) |>
#   rename(c("tSNE1" = "c_x",
#            "tSNE2" = "c_y"))
# 
# scaled_mnist_data_model <- bind_cols(scaled_mnist_data_model, model_2d)

# Combine with the true model for visualization
# df <- dplyr::bind_rows(scaled_mnist_data_model |> mutate(type = "model"),
#                        scaled_mnist_data)

# scaled_mnist_data <- scaled_mnist_data |>
#   select(-type)

@@

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  #dplyr::mutate(type = df_exe$type) |>
  dplyr::mutate(ID = dplyr::row_number()) |>
  mutate(cluster = tsne_minst_scaled_temp$cluster)

@@

mnist_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.2,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000","#e41a1c")) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("b") +
  theme(
    legend.position = "none"
  )

@@

mnist_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.2,
    alpha = 0.1) +
  scale_color_manual(values = c("#000000", "#e41a1c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c") +
  theme(
    legend.position = "none"
  )

@@

mnist_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.2,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#e41a1c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("d") +
  theme(
    legend.position = "none"
  )
